CudeleFS: Programmable Consistency and \\Durability in a Global Namespace

### 2017-04-02 paper/sections/1_introduction.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
\caption{Administrators can assign weaker consistency and durability policies
to subtrees to get the same performance benefits of state-of-the-art HPC
architectures. Applications that need stronger guarantees can still reside in
the same namespace.
}\label{fig:subtree-policies}
\end{figure}

@@

% What did we do
We propose a subtree policy API that lets future programmers control
the consistency and durability for subtrees in the file system namespace. For performance, one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  Cudele supports 3 forms of consistency
(invisible, weak, and strong) and 3 degrees of durability (none, local, and
global) giving the administrator a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

@@

  \item a prototype that lets administrators program a range of
  consistency and durability semantics (9 permutations), allowing them to custom
  fit the storage system to the application.

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.
Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
prototype and API. Section~\ref{sec:implementation} describes Cudele's
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places Cudele in the context of other related
work.

>>>>

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
\caption{Users can assign weaker consistency and durability policies
to subtrees to get the same performance benefits of state-of-the-art HPC
architectures. Applications that need stronger guarantees can still reside in
the same namespace.
}\label{fig:subtree-policies}
\end{figure}

@@

% What did we do
We propose a subtree policy API that lets future programmers control
the consistency and durability for subtrees in the file system namespace. For performance, one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  CudeleFS supports 3 forms of consistency
(invisible, weak, and strong) and 3 degrees of durability (none, local, and
global) giving the user a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

@@

  \item a prototype that lets users program a range of
  consistency and durability semantics (9 permutations), allowing them to custom
  fit the storage system to the application.

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.
Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
prototype and API. Section~\ref{sec:implementation} describes Cudele's
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places CudeleFS in the context of other related
work.

### 2017-04-02 paper/sections/4_methodology.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

In this section we describe CudeleFS, our file system that lets administrators
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. CudeleFS exposes these
mechanisms and the administrator composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

@@

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Create"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisified by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes an in-memory journal on the
client and applies the updates directly to the in-memory namespace maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- CudeleFS makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

@@

%The metadata store and journal
%are different ways of representing the namespace.  Cudele presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Create" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
mechanisms used by default in CephFS.  Using existing configuration settings in
Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
will not save journals in the object store and the daemons that apply the
journal to the metadata store will never run.  For ``Local Persist", clients
write serialized log events to a file on local disk and for ``Global Persist",
clients push the journal into the objects store. The overheads for both ``Local
Persist" and ``Global Persist" is the write bandwidth of the local disk and
object store, respectively.  These persist mechanisms are part a library that
links into the client.

@@

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create"
    "durability": "local_persist"
}
\end{minted}
\caption{Implementing DeltaFS with CudeleFS.}
\label{src:deltafs}
\end{listing}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create+volatile_apply"
    "durability": "local_persist"
}
\end{minted}
\caption{Implementing BatchFS with CudeleFS.}
\label{src:batchfs}
\end{listing}

% describe table
The spectrum of consistency and durability guarantees that administrators can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``weak" merges updates at some time in the
future ({\it e.g.}, when the system has time, when number of updates reaches a
certain threshold, when the client is done writing, etc.); and updates in
``strong" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability; DeltaFS uses ``invisible" consistency
and ``local" durability and BatchFS uses ``weak" consistency and ``local"
durability.  To compose the mechanisms administrators inject which mechanisms
(described in Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use
in parallel using a domain specific language.

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both mechanisms do the same
thing or \texttt{stream+save} since ``global" durability is stronger and has
more overhead than ``local" durability.

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed globally consistent until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that changes a subtree from ``local
persist" to ``global persist", CudeleFS no guarantee until the mechanisms are
complete.

@@

\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

\subsection{Cudele Namespace API}
\label{sec:cudele-namespace-api}

% the interface
The interface for setting the subtree policies is with \{path,
\(<\)block\(|\)overwrite\(>\), pre-allocated inodes\} tuples. For example:

\texttt{(msevilla/mydir, policies.yml)}

would decouple the path \texttt{msevilla/mydir} and would apply the policies in
\texttt{policies.yml}. The policies file supports the following values:

@@

  \item \texttt{allocated\_inodes}: the number of inodes to allocate to the
  decoupled namespace (default 100)

  \item \texttt{interfere\_callback}: how to handle a request from another
  client targeted at the now decoupled subtree (default \texttt{overwrite})

  \item \texttt{consistency\_callback}: which consistency model to use (default
  \texttt{RPCs})

  \item \texttt{durability\_callback}: which durability model to use (default
  \texttt{stream})

@@

% description
For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources
for updates that may get overwritten. If the application does not mind losing
updates, for example it wants approximations for results that take too long to
compute, it can select \texttt{overwrite}. In this case, metadata will be
written and the computation from the decoupled namespace will take priority at
merge time because the results are more accurate.

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
would use the configuration from Listing~\ref{src:deltafs}. BatchFS's merging
back into the global namespace can be achieved with the configuration in
Listing~\ref{src:batchfs}.

%\begin{listing}
%\begin{minted}[frame=single,
%               framesep=3mm,
%               xleftmargin=21pt,
%               tabsize=4]{js}
%{     
%    "allocated_inodes": "-"
%    "interfere_policy": "-"
%    "consistency": "RPCs"
%    "durability": "stream"
%}
%\end{minted}
%\caption{Existing CephFS on Cudele.}
%\label{src:batchfs}

%\end{listing}

>>>>

In this section we describe CudeleFS, our file system that lets users
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. CudeleFS exposes these
mechanisms and the user composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

@@

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Create"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisified by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
client and applies the updates directly to the in-memory metadata store maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- CudeleFS makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

@@

%The metadata store and journal
%are different ways of representing the namespace.  CudeleFS presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Create" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
mechanisms used by default in CephFS.  Using existing configuration settings in
Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
will not save journals in the object store. For ``Local Persist", clients write
serialized log events to a file on local disk and for ``Global Persist",
clients push the journal into the objects store. The overheads for both ``Local
Persist" and ``Global Persist" is the write bandwidth of the local disk and
object store, respectively.  These persist mechanisms are part of the library that
links into the client.

@@

@@

\caption{Users can explore the consistency (C) and
durability (D) spectrums by composing CudeleFS mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that users can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``weak" merges updates at some time in the
future ({\it e.g.}, when the system has time, when the number of updates reaches a
certain threshold, when the client is done writing, etc.); and updates in
``strong" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability ~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{creates+RPCs} since
both mechanisms do the same thing or \texttt{stream+save} since ``global"
durability is stronger and has more overhead than ``local" durability.

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed to have ``global" durability until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that changes a subtree from ``local
persist" to ``global persist", CudeleFS makes no guarantee until the mechanisms are
complete.

\subsection{CudeleFS Namespace API}
\label{sec:cudelefs-namespace-api}

% the interface
Users control consistency and durability for subtrees by contacting the monitor
daemon with a directory path and presesenting a policies configuration file.
For example, (msevilla/mydir, policies.yml) would decouple the path
``msevilla/mydir" and would apply the policies in ``policies.yml".  The
policies file supports the following parameters (default values are in parenthesis):

@@

  \item \texttt{consistency}: which consistency model to use (\texttt{RPCs})

  \item \texttt{durability}: which durability model to use (\texttt{stream})

  \item \texttt{allocated\_inodes}: the number of inodes to provision to the
  decoupled namespace (100)

  \item \texttt{interfere\_policy}: how to handle a request from another
  client targeted at this subtree (\texttt{allow}))

@@

The ``Consistency" and ``Durability" parameters are set to the CudeleFS
mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
``Allocated Inodes" is a way for the application to specify how many files it
intends to create. It is a contract so that the file system can provision
enough resources for the incumbent merge and so it can give valid inodes to
other clients. The inodes can be used anywhere within the decoupled namespace
({i.e.} at any depth in the subtree).

% description
``Interfere Policy" has two settings: \texttt{block} and \texttt{allow}.
For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources
for updates that may get overwritten. If the application does not mind losing
updates, for example it wants approximations for results that take too long to
compute, it can select \texttt{allow}. In this case, metadata from the interferring client will be
written and the computation from the decoupled namespace will take priority at
merge time because the results are more accurate.

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
would use the configuration:

\begin{listing}[h]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

BatchFS merges updates into the global namespace after the job is complete.
This can be achieved with by composing two mechanisms:

\begin{listing}[h]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create+volatile_apply"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

### 2017-04-02 paper/sections/5_implementation.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% how i used it
\noindent\textbf{CudeleFS}: uses the metadata store format to write objects
formatted like the journal into the object store. By writing with the same
format, the metadata servers can read and use the recovery code to materialize
the updates from a client's decoupled namespace ({\it i.e.} merge).  These
implementations required no changes to CephFS because the metadata servers know
how to read the events the library is writing into the object store.  By
re-using the journal subsystem to implement the namespace decoupling, Cudele
leverages the write/read optimized data structures, the formats for persisting
events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
functions for replaying events onto the internal namespace data structures.\\

\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply

@@

\noindent\textbf{CudeleFS}: uses the journal subsystems within the metadata
server to replay updates onto the in-memory metadata store.  When the clients
are ready to merge their updates back into the global namespace, they pass a
binary file of  metadata updates to the metadata server. The metadata server
uses its recovery code to ``replay" the updates.\\

\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply, Local/Persist

@@

\noindent\textbf{CudeleFS}: using the journal tool, clients can materialize the
journal in memory and append events to their own copy. When they are done they
export their journal to a binary file. Clients also use the journal tool to
save their updates to disk or to the object store.\\

\noindent\textbf{Used to Implement}: Create, Volatile/Nonvolatile Apply

@@

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing operations like \texttt{lookup()}.  For example, if a client or
metadata server is not caching the directory inode, all creates within that
directory will result in a lookup and a create request. If the directory inode
is cached then only the create needs to be sent.  The size of the inode cache
is configurable so as not to saturate the memory on the metadata server --
inodes in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.\\

\noindent\textbf{Used to Implement}: Nonvolatile Apply, Create

@@

In CephFS, inodes store policies for policies, like how the file is striped
across the object store or for managing subtrees for load balancing. These
policies adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other.  \\

\noindent\textbf{Cudele}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

>>>>

% how i used it
\noindent\textbf{CudeleFS}: the ``RPCs" mechanism uses the metadata store to
service requests. Using code designed for recovery, ``Nonvolatile Apply" and
``Volatile Apply" replay updates onto the metadata store in memory and in the
object store, respectively.  When the clients are ready to merge their updates
back into the global namespace, they pass a binary file of  metadata updates to
the metadata server.

%\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply

@@

\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Create",
``Local Persist", and ``Global Persist".  ``Stream" is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemeted by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, CudeleFS leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

@@

\noindent\textbf{CudeleFS}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
``Create", ``Volatile Apply", and ``Nonvolatile Apply".

@@

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For example, if a client or
metadata server is not caching the directory inode, all creates within that
directory will result in a lookup and a create request. If the directory inode
is cached then only the create needs to be sent.  The size of the inode cache
is configurable so as not to saturate the memory on the metadata server --
inodes in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

\noindent\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
code to allocate inodes to clients that decouple parts of the namespace and to
skip inodes used by the client at merge time.\\

@@

In CephFS, inodes already store policies, like how the file is striped
across the object store or for managing subtrees for load balancing. These
policies adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other.  \\

\noindent\textbf{Cudele}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different CudeleFS mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

### 2017-04-02 paper/sections/6_results.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

We evaluate Cudele on a 15 node cluster, partitioned into 8 object storage
servers, 3 metadata servers, and 2 monitor servers. The object storage servers
double as clients which is fine because clients are CPU and memory bound while
object storage servers are disk IO bound. All daemons run as a single process
which is the default setting for Ceph and the nodes have 2 dual core 2GHz
processors with 8GB of RAM. There are three daemons per object storage server
(one for each disk formatted with XFS) and they share an SSD for the journal.
The nodes are running Ubuntu 12.04.4, kernel version 3.2.0-63 but all
experiments run in Docker containers; this makes it easier to tear down and
re-initialize ({\it e.g.}, dropping the kernel cache) for the cluster between
experiments.

@@

\subsection{Cudele Mechanism Performance}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
The performance of the Cudele mechanisms normalized to the runtime of the
create mechanism. The runtime of the create mechanism is the time it takes to
write 100 file creates to the client's in-memory journal of metadata
updates.\label{fig:slowdown-mechanisms}}
\end{figure}

@@

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
\subsubsection{Overhead of ``nonvolatile apply"} The cost of ``nonvolatile
apply" is much larger than all the other mechanisms.  That mechanism was not
implemented as part of Cudele -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communcating through the object store is shown
by comparing the runtime of ``volatile apply" + ``global persist" to
``nonvolatile apply". These two operations end up with the same final metadata
state but using ``nonvolatile apply" is clearly inferior.

@@

% Final file system metadata states are equivalent No guarantees while
% transitioning mechanisms
We compare these semantics because the final metadata states are equivalent.
Cudele makes no guarantees during execution of the mechanisms or when
transitioning semantics -- the semantics are guaranteed {\it once the mechanism
completes}. So if servers fail during a mechanism, metadata or data may be
lost.

@@

\begin{listing}[tb]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\caption{Existing CephFS on Cudele with blocking.}
\label{src:block}
\end{listing}

>>>>

We evaluate CudeleFS on a 15 node cluster, partitioned into 8 object storage
servers, 3 metadata servers, and 2 monitor servers. The object storage servers
double as clients which is fine because clients are CPU and memory bound while
object storage servers are disk IO bound. All daemons run as a single process
which is the default setting for Ceph and the nodes have 2 dual core 2GHz
processors with 8GB of RAM. There are three daemons per object storage server
(one for each disk formatted with XFS) and they share an SSD for the journal.
The nodes are running Ubuntu 12.04.4, kernel version 3.2.0-63 but all
experiments run in Docker containers; this makes it easier to tear down and
re-initialize ({\it e.g.}, dropping the kernel cache) for the cluster between
experiments.

@@

\subsection{CudeleFS Mechanism Performance}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
The performance of the CudeleFS mechanisms normalized to the runtime of the
create mechanism. The runtime of the create mechanism is the time it takes to
write 100 file creates to the client's in-memory journal of metadata
updates.\label{fig:slowdown-mechanisms}}
\end{figure}

@@

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
\subsubsection{Overhead of ``nonvolatile apply"} The cost of ``nonvolatile
apply" is much larger than all the other mechanisms.  That mechanism was not
implemented as part of CudeleFS -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communcating through the object store is shown
by comparing the runtime of ``volatile apply" + ``global persist" to
``nonvolatile apply". These two operations end up with the same final metadata
state but using ``nonvolatile apply" is clearly inferior.

@@

% Final file system metadata states are equivalent No guarantees while
% transitioning mechanisms
We compare these semantics because the final metadata states are equivalent.
CudeleFS makes no guarantees during execution of the mechanisms or when
transitioning semantics -- the semantics are guaranteed {\it once the mechanism
completes}. So if servers fail during a mechanism, metadata or data may be
lost.

@@

\begin{listing}[tb]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\caption{Existing CephFS on CudeleFS with blocking.}
\label{src:block}
\end{listing}

### 2017-04-02 paper/sections/7_conclusion.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% How awesome programmability is
``Programmability" is a way of designing storage systems that encourages code
re-use and composition.  A programmable storage system exposes internal
subsystem as building blocks for higher level services. This `dirty-slate'
approach limits redundant code and leverages the robustness of the underlying
storage system. Cudele uses this approach and re-uses some of the building
blocks from the Malacology programmable storage system~\cite{sevilla:eurosys17}
to great success and requires only:

@@

Relaxing consistency and durability semantics in the file system is a
double-edged sword. While it achieves better performance and scalability, it
alienates applications that rely on strong consistency and durability. CudeleFS
lets administrators assign consistency and durability guarantees to subtrees in
the global namespace, giving administrators a wide range of policies
optimizations that can be custom fit to the application. Using CudeleFS, we
compare related work on the same file system and conclude that strong
consistency can cause a 104\(\times\) slow down while merging updates and
maintaining durability only have between a \(7-10\times\) slow down.

>>>>

% How awesome programmability is
``Programmability" is a way of designing storage systems that encourages code
re-use and composition.  A programmable storage system exposes internal
subsystem as building blocks for higher level services. This `dirty-slate'
approach limits redundant code and leverages the robustness of the underlying
storage system. CudeleFS uses this approach and re-uses some of the building
blocks from the Malacology programmable storage system~\cite{sevilla:eurosys17}
to great success and requires only:

@@

Relaxing consistency and durability semantics in the file system is a
double-edged sword. While it achieves better performance and scalability, it
alienates applications that rely on strong consistency and durability. CudeleFS
lets users assign consistency and durability guarantees to subtrees in
the global namespace, giving users a wide range of policies
optimizations that can be custom fit to the application. Using CudeleFS, we
compare related work on the same file system and conclude that strong
consistency can cause a 104\(\times\) slow down while merging updates and
maintaining durability only have between a \(7-10\times\) slow down.
